Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-04-17 paper/arm.tex
no dangling refs

\begin{theorem}
  \label{thm:ec}
  Suppose $\aEG_1$ is $({\rco_1}, {\rrfx_1}, {\rgcb_1})$-valid for $\aCmd$
  under the suboptimal lowering that maps non-relaxed reads to
  $(\DMBSY\SEMI\LDAR)$.  Then there is a top-level pomset
  $\aPS_2\in\semrr{\aCmd}$ such that $\aEvs_2=\aEvs_1$,
  $\labeling_2=\labeling_1$, ${\rrfx_2}={\rrfx_1}$, and ${\le_2}={\rgcb_1}$.

@@

    \vspace{-.5\baselineskip}
    \begin{lemma}
      \label{lemma:fr}
      Suppose $\aEG$ is $({\rco}, {\rrfx}, {\rgcb})$-valid.  Then
      ${\rgcb}\supseteq{\rfr}$.

@@

      \vspace{-.5\baselineskip}
      \begin{proof}
        By way of contradiction, assume $\aEv\xgcb\cEv$.  If $\cEv\xrfx\aEv$
        then by \ref{arm-gcb} we must also have $\cEv\xgcb\aEv$,
        contradicting the assumption that $\rgcb$ is a total order.
        %
        Otherwise that there is some $\bEv\neq\cEv$ such that
        $\bEv\xrfx\aEv$, and therefore $\bEv\xgcb\aEv$.  By transitivity,
        $\bEv\xgcb\cEv$.  By the definition of $\rfr$, we have
        $\aEv\xfr\cEv$.  But this contradicts \ref{arm-local}, since
        $\cEv\xpoloc\aEv$.
        %Applying \ref{arm-gcb-blocks}, we have that either $\cEv\xgcb\bEv$ or $\aEv\xgcb\cEv$.
      \end{proof}
    \end{lemma}
    We show that all the order required in the pomset is also required by
    \armeight{}.  \ref{rf-block} holds since ${\rcb_1}$ is consistent with
    ${\rco_1}$ and ${\rfr_1}$.  As noted \ref{lob-le}, $\rlob$ includes the order
    required by \ref{seq-kappa} and \ref{seq-le-delays}.  We need only show
    that the order removed from \ref{arm-gcb-lob} can also be removed from
    the pomset.  In order for \label{arm-gcb-lob} to remove order from $\aEv$
    to $\cEv$, we must have $\bEv\xrfx\aEv$ and $\bEv\xpoloc\aEv$ but not
    $\bEv\xlob\cEv$.  Because of our suboptimal lowering, it must be that
    $\aEv$ is a relaxed read; otherwise the $\DMBSY$ would require
    $\bEv\xlob\cEv$.  Thus we know that \ref{seq-le-delays} does not require
    order from $\aEv$ to $\cEv$.  By chaining \ref{read-tau-ind} and
    \ref{write-tau}, any dependence on the read can by satisfied without
    introducing order in \ref{seq-kappa}.
  \end{proof}  
\end{theorem}

@@

The generality of \refthm{thm:ec} is not limited by the assumption that
${\rcb_1}\supseteq{\rfr_1}$:
\begin{lemma}
  \label{lemma:fr}
  Suppose $\aEG$ is \EC-valid via $({\rco}, {\rrfx}, {\rcb})$.  Then there a
  permutation ${\rcbp}$ of ${\rcb}$ such that $\aEG$ is \EC-valid via
  $({\rco}, {\rrfx}, {\rcbp})$ and ${\rcbp}\supseteq{\rfr}$, where ${\rfr}$
  is defined in \ref{arm-local}.

>>>>

\begin{theorem}
  \label{thm:egc}
  Suppose $\aEG_1$ is $({\rco_1}, {\rrfx_1}, {\rgcb_1})$-valid for $\aCmd$
  under the suboptimal lowering that maps non-relaxed reads to
  $(\DMBSY\SEMI\LDAR)$.  Then there is a top-level pomset
  $\aPS_2\in\semrr{\aCmd}$ such that $\aEvs_2=\aEvs_1$,
  $\labeling_2=\labeling_1$, ${\rrfx_2}={\rrfx_1}$, and ${\le_2}={\rgcb_1}$.

@@

    \vspace{-.5\baselineskip}
    \begin{lemma}
      \label{lemma:fr1}
      Suppose $\aEG$ is $({\rco}, {\rrfx}, {\rgcb})$-valid.  Then
      ${\rgcb}\supseteq{\rfr}$.

@@

      \vspace{-.5\baselineskip}
      \begin{proof}
        By way of contradiction, assume $\aEv\xgcb\cEv$.  If $\cEv\xrfx\aEv$
        then by \ref{arm-gcb} we must also have $\cEv\xgcb\aEv$,
        contradicting the assumption that $\rgcb$ is a total order.
        %
        Otherwise that there is some $\bEv\neq\cEv$ such that
        $\bEv\xrfx\aEv$, and therefore $\bEv\xgcb\aEv$.  By transitivity,
        $\bEv\xgcb\cEv$.  By the definition of $\rfr$, we have
        $\aEv\xfr\cEv$.  But this contradicts \ref{arm-local}, since
        $\cEv\xpoloc\aEv$.
        %Applying \ref{arm-gcb-blocks}, we have that either $\cEv\xgcb\bEv$ or $\aEv\xgcb\cEv$.
      \end{proof}
    \end{lemma}
    We show that all the order required in the pomset is also required by
    \armeight{}.  \ref{rf-block} holds since ${\rcb_1}$ is consistent with
    ${\rco_1}$ and ${\rfr_1}$.  As noted \ref{lob-le}, $\rlob$ includes the order
    required by \ref{seq-kappa} and \ref{seq-le-delays}.  We need only show
    that the order removed from \ref{arm-gcb-lob} can also be removed from
    the pomset.  In order for \ref{arm-gcb-lob} to remove order from $\aEv$
    to $\cEv$, we must have $\bEv\xrfx\aEv$ and $\bEv\xpoloc\aEv$ but not
    $\bEv\xlob\cEv$.  Because of our suboptimal lowering, it must be that
    $\aEv$ is a relaxed read; otherwise the $\DMBSY$ would require
    $\bEv\xlob\cEv$.  Thus we know that \ref{seq-le-delays} does not require
    order from $\aEv$ to $\cEv$.  By chaining \ref{read-tau-ind} and
    \ref{write-tau}, any dependence on the read can by satisfied without
    introducing order in \ref{seq-kappa}.
  \end{proof}  
\end{theorem}

@@

The generality of \refthm{thm:ec} is not limited by the assumption that
${\rcb_1}\supseteq{\rfr_1}$:
\begin{lemma}
  \label{lemma:fr2}
  Suppose $\aEG$ is \EC-valid via $({\rco}, {\rrfx}, {\rcb})$.  Then there a
  permutation ${\rcbp}$ of ${\rcb}$ such that $\aEG$ is \EC-valid via
  $({\rco}, {\rrfx}, {\rcbp})$ and ${\rcbp}\supseteq{\rfr}$, where ${\rfr}$
  is defined in \ref{arm-local}.
